---
title: "Anhedonia in PTSD and MDD"
author: "Or Duek"
output:
  html_document:
    df_print: paged
---

## Assessing anhedonia in PTSD

```{r, echo=FALSE, results='hide', comment=FALSE, warning=FALSE, message=FALSE}
require(tidyverse)
require(readxl)
require(ggplot2)
require(lme4)
require(lmerTest)
require(ppcor)
require(lsr)
require(heplots)
```

```{r, echo=FALSE}
# load data
df <- read_csv('/home/or/PTSD_reward_data/CleanedOfficialDataSet11_23_2020.csv')
table(df$`CAPS PTSD`, df$scid_mdd_pm)
# adding caps without D
df$caps_noD <- df$`CAPS Total` -  df$`Crit D Total`
df$caps_noEN <- df$`CAPS Total` - df$EN
```

## Descriptive statistics
- Need to fix the data set for that

```{r}
# sex
table(df$Sex)
77 / (77+65) # 54% female
summary(df$Age)
```

```{r}
# Number diagnosed with only PTSD
table(df$scid_mdd_pm, df$`CAPS PTSD`)

onlyPTSD <- dplyr::filter(df, df$`CAPS PTSD`=='PTSD' & df$scid_mdd_pm == 'No MDD')
onlyMDD <- dplyr::filter(df, df$`CAPS PTSD`=='NO PTSD' & df$scid_mdd_pm == 'Yes MDD')
bothMDD_PTSD <- dplyr::filter(df, df$`CAPS PTSD`=='PTSD' & df$scid_mdd_pm == 'Yes MDD')
HC <- dplyr::filter(df, df$`CAPS PTSD`=='NO PTSD' & df$scid_mdd_pm == 'No MDD')

# PTSD
sd(onlyPTSD$Age)
table(onlyPTSD$Sex)
mean(onlyPTSD$`CAPS Total`)
sd(onlyPTSD$`CAPS Total`)
# MDD
mean(onlyMDD$Age) # 35.25
sd(onlyMDD$Age) # 10.09
table(onlyMDD$Sex) # 5f/3m
mean(onlyMDD$`CAPS Total`) # 22.25
sd(onlyMDD$`CAPS Total`) # 11.93
# Both
mean(bothMDD_PTSD$Age) # 40.98
sd(bothMDD_PTSD$Age) # 12.85
table(bothMDD_PTSD$Sex) # 36f/24m
mean(bothMDD_PTSD$`CAPS Total`) # 44.88
sd(bothMDD_PTSD$`CAPS Total`) # 10.13

# None
mean(HC$Age) # 39.53
sd(HC$Age) # 13.47
table(HC$Sex) # 8f/18m
mean(HC$`CAPS Total`) # 11.25
sd(HC$`CAPS Total`) # 8.39

```

### First look at overall symtpoms


```{r, echo=FALSE}
ggplot(df, aes(y=`Crit D Total`, x= scid_mdd_pm, color=`CAPS PTSD`)) + geom_boxplot() + geom_jitter() + xlab("MDD") + ylab("Cluster D Total Score") + theme_minimal()
```

* This shows nicely that PTSD patients score higher in the anhedonia scores, even when not depressed
* Lets tests it using anova



```{r}
model1 <- aov(`Crit D Total` ~ scid_mdd_pm*`CAPS PTSD`, data=df)
summary(model1)

#eta1 <- etasq(model1, partial = FALSE) # don't trust this function
# calculate manually
#model1

PTSD_parEta <- 1581.2754 / (1581.2754+2367.4913) # 0.4
PTSD_parEta
mdd_parEta <- 2044.7645 / (2044.7645 + 2367.4913) # 0.46
mdd_parEta
```

- As expected, we see a significant effect for both MDD and PTSD.
- 40% partial eta square for the PTSD diagnosis, while only 31% for depression. (in different calculation we have 46% explained by depression). 
- Lets compare the model with one with only depression

```{r}
model1_1 <- aov(`Crit D Total` ~ scid_mdd_pm, data=df)
summary(model1_1)
anova(model1, model1_1)
BIC(model1, model1_1)

```
- So, model including PTSD significantly fits better than model with only depression. 
- Lets use PCL (which is continuous and we don't use the exact same measure)

## Gender interaction

```{r}
modelG <- aov(`Crit D Total` ~ scid_mdd_pm*Sex+`CAPS PTSD`*Sex, data=df)
summary(modelG)

```


### Using BDI
- Lets take a look only on the PTSD group and see how it behaves


```{r}
# test bdi on entire population
df_valid_bdi_all <- df[!is.na(df$`Scr BDI`), ] # 64 with valid BDI scores
df_valid_bdi_all$bdi <- as.numeric(df_valid_bdi_all$`Scr BDI`)
model3 <- lm(`Crit D Total` ~ `CAPS PTSD` * bdi, df_valid_bdi_all)
summary(model3)

eta3 <- etasq(model3) # most variance explained by bdi, but almost 10% by PTSD
```

- Lets try with PCL also

```{r}
df_valid_bdi_all$pcl <- as.numeric(df_valid_bdi_all$`Scr PCL`)
df_valid_bdi_all <- df_valid_bdi_all[!is.na(df_valid_bdi_all$`Scr PCL`), ] # 57

model4 <- lm(`Crit D Total` ~ bdi+pcl, data=df_valid_bdi_all)
summary(model4)
BIC(model4)
eta4 <- etasq(model4) # here pcl explains almost 9%
```

- Compare to model with only BDI

```{r}
model4.2 <- lm(`Crit D Total` ~ bdi, data=df_valid_bdi_all)
BIC(model4, model4.2)
anova(model4, model4.2)
```

## Now we use the total CAPS score without D as a continous variable (with BDI)

```{r}
df$caps_noD <- df$`CAPS Total` -  df$`Crit D Total`
model5 <- lm(`Crit D Total` ~ bdi*caps_noD, data=df_valid_bdi_all)
summary(model5)

eta5 <- etasq(model5) # here pcl explains almost 9%
```

- Compare to model without CAPS

```{r}
model5.2 <- lm(`Crit D Total` ~ bdi, data=df_valid_bdi_all)
BIC(model5, model5.2)
anova(model5, model5.2)
```

- Caps score continous without the D criteria explains almost 30 percent of variance. 
#####################################################################################################
####################################################################################################

As Robb suggested - testing overall severity:
```{r}
modelS <- aov(`CAPS Total` ~ scid_mdd_pm, data=df)
table(df$`CAPS Total`, df$scid_mdd_pm)
summary(modelS)
```


# These are results for the paper:

## Now lets do the same with Emotional numbing symptoms 

- The paper might focus primarily on that element.

```{r}
pdf('Figure_1.pdf')
ggplot(df, aes(y=EN, x= scid_mdd_pm, color=`CAPS PTSD`)) + geom_boxplot() + geom_jitter() + xlab("MDD") + ylab("Emotional Numbing Score") + theme_minimal()
dev.off()
```


```{r}
model5 <- aov(EN ~ scid_mdd_pm * `CAPS PTSD`, data=df)
summary(model5)

eta5 <- etasq(model5, partial = T) # don't trust this function
# calculate manually
#model1

PTSD_parEta <- 273.6 / (273.6+853.3) # 0.24
PTSD_parEta
mdd_parEta <- 573.4 / (573.4 + 853.3) # 0.4
mdd_parEta
```

- Test only depression to compare models

```{r}
model5_1 <- aov(EN ~ scid_mdd_pm, data=df)
summary(model5_1)
eta5_1 <- etasq(model5_1, partial = T)
anova(model5, model5_1)
BIC(model5, model5_1)
```
- Advantage to the model including both MDD and PTSD

```{r}
df_valid_bdi_all$caps_noEN <- df_valid_bdi_all$`CAPS Total` -  df_valid_bdi_all$EN
model6 <- lm(EN ~ bdi + caps_noEN, data=df_valid_bdi_all)
summary(model6)

eta6 <- etasq(model6) # 9%
```

- Compare to model with just BDI

```{r}
model6.2 <- lm(EN ~ bdi, data=df_valid_bdi_all)
summary(model6.2)
eta6.2 <- etasq(model6.2)
BIC(model6, model6.2, model6.3)
anova(model6, model6.2) # sig. difference, by some points...
```

- Compare with model of just CAPS

```{r}
model6.3 <- lm(EN ~ caps_noEN, data=df_valid_bdi_all)
summary(model6.3)
eta6.3 <- etasq(model6.3)
BIC(model6, model6.2, model6.3)
anova(model6.2, model6.3)#, model6.3) # sig. difference, by some points...
```

## Test with RCF data

```{r}
dfRCF <- read.csv('/home/or/rewardPTSD/ReconsolidationOfFea_DATA_2021-03-02_1737.csv')

table(dfRCF$caps_ptsdc)
table(dfRCF$scid_mdd_pm)
table(dfRCF$caps_ptsdc, dfRCF$scid_mdd_pm)
```

```{r}
dfBDI <- dplyr::select(dfRCF, starts_with('bdi'), -bdiii_complete)
bdiTotal <- rowSums(dfBDI)
dfRCF$bdi_total <- bdiTotal

dfRCF_calc <- dplyr::select(dfRCF, subject_id,scid_mdd_pm, caps_ptsdc, caps_c4_im, caps_c5_im, caps_c6_im, caps_c4_fm,
  caps_c5_fm, caps_c6_fm, capstotalwithouten, calculatedtotal)
dfRCF_calc <- na.omit(dfRCF_calc)
dfRCF_calc$EN <- dfRCF_calc$calculatedtotal - dfRCF_calc$capstotalwithouten
dfRCF_calc$MDD <- case_when(dfRCF_calc$scid_mdd_pm == 3 ~ "YES", TRUE ~ "No")
dfRCF_calc$PTSD <- case_when(dfRCF_calc$caps_ptsdc == 1 ~ "YES", TRUE ~ "No")

# add bdi
bdiDF <- dplyr::select(dfRCF, subject_id, bdi_total)
bdiDF <- na.omit(bdiDF)
hist(bdiDF$bdi_total)
summary(bdiDF$bdi_total)
dfRCF_calc <- merge(dfRCF_calc, bdiDF, by = 'subject_id', all.x = T)
```

Do descriptive

```{r}
table(dfRCF$dem_gender)
table(dfRCF_calc$MDD)
table(dfRCF_calc$PTSD)
table(dfRCF_calc$MDD, dfRCF_calc$PTSD)
```

Run analysis

```{r}
model1_RCF <- aov(EN ~ MDD * PTSD, data=dfRCF_calc)
summary(model1_RCF)

etasq(model1_RCF) # here pcl explains almost 9%
pdf('FigureS1.pdf')
ggplot(dfRCF_calc, aes(x=MDD, y = EN,  color = PTSD)) + geom_boxplot() + geom_jitter() + ylab("Emotional Numbing Score") + theme_minimal()
dev.off()
```

Test model with just mdd

```{r}
model1.2RCF <- aov(EN ~ MDD, data=dfRCF_calc)
summary(model1.2RCF)
etasq(model1.2RCF)
BIC(model1_RCF, model1.2RCF)
anova(model1_RCF, model1.2RCF)
```

- Now do linear regression

```{r}
model2.RCF <- lm(EN ~ capstotalwithouten * bdi_total, data = dfRCF_calc)
summary(model2.RCF)
etasq(model2.RCF)
```

Model with just BDI

```{r}
model2.2RCF <- lm(EN ~ bdi_total, data = dfRCF_calc)
summary(model2.2RCF)
etasq(model2.2RCF)

BIC(model2.RCF, model2.2RCF)
anova(model2.RCF, model2.2RCF)
```
```{r}
dfRCF_age <- read.csv('/home/or/rewardPTSD/ReconsolidationOfFea_DATA_2021-03-03_0913.csv')
na.omit(dfRCF_age)
summary(dfRCF_age$dem_age)
sd(dfRCF_age$dem_age)
```

