---
title: "Anhedonia in PTSD and MDD"
author: "Or Duek"
output:
  html_document:
    df_print: paged
---

## Assessing anhedonia in PTSD

```{r, echo=FALSE, results='hide', comment=FALSE, warning=FALSE, message=FALSE}
require(tidyverse)
require(readxl)
require(ggplot2)
require(lme4)
require(lmerTest)
require(ppcor)
require(lsr)
require(heplots)
library(relaimpo)
```

```{r, echo=FALSE}
# load data
df <- read_csv('/home/or/PTSD_reward_data/CleanedOfficialDataSet11_23_2020.csv')
table(df$`CAPS PTSD`, df$scid_mdd_pm)
# adding caps without D
df$caps_noD <- df$`CAPS Total` -  df$`Crit D Total`
df$caps_noEN <- df$`CAPS Total` - df$EN
```

## Descriptive statistics
- Need to fix the data set for that

```{r}
# sex
table(df$Sex)
77 / (77+65) # 54% female
summary(df$Age)
```

```{r}
# Number diagnosed with only PTSD
table(df$scid_mdd_pm, df$`CAPS PTSD`)

onlyPTSD <- dplyr::filter(df, df$`CAPS PTSD`=='PTSD' & df$scid_mdd_pm == 'No MDD')
onlyMDD <- dplyr::filter(df, df$`CAPS PTSD`=='NO PTSD' & df$scid_mdd_pm == 'Yes MDD')
bothMDD_PTSD <- dplyr::filter(df, df$`CAPS PTSD`=='PTSD' & df$scid_mdd_pm == 'Yes MDD')
HC <- dplyr::filter(df, df$`CAPS PTSD`=='NO PTSD' & df$scid_mdd_pm == 'No MDD')

# PTSD
sd(onlyPTSD$Age)
table(onlyPTSD$Sex)
mean(onlyPTSD$`CAPS Total`)
sd(onlyPTSD$`CAPS Total`)
# MDD
mean(onlyMDD$Age) # 35.25
sd(onlyMDD$Age) # 10.09
table(onlyMDD$Sex) # 5f/3m
mean(onlyMDD$`CAPS Total`) # 22.25
sd(onlyMDD$`CAPS Total`) # 11.93
# Both
mean(bothMDD_PTSD$Age) # 40.98
sd(bothMDD_PTSD$Age) # 12.85
table(bothMDD_PTSD$Sex) # 36f/24m
mean(bothMDD_PTSD$`CAPS Total`) # 44.88
sd(bothMDD_PTSD$`CAPS Total`) # 10.13

# None
mean(HC$Age) # 39.53
sd(HC$Age) # 13.47
table(HC$Sex) # 8f/18m
mean(HC$`CAPS Total`) # 11.25
sd(HC$`CAPS Total`) # 8.39

```
#####################################################################################################
####################################################################################################

## Emotional numbing symptoms 

- The paper might focus primarily on that element.

```{r}
pdf('Figure_1.pdf')
ggplot(df, aes(y=EN, x= scid_mdd_pm, color=`CAPS PTSD`)) + geom_boxplot() + geom_jitter() + xlab("MDD") + ylab("Emotional Numbing Score") + theme_minimal()
dev.off()
```


```{r}
model5 <- aov(EN ~ scid_mdd_pm * `CAPS PTSD`, data=df)
summary(model5)

eta5 <- etasq(model5, partial = T) # don't trust this function
# calculate manually
#model1

PTSD_parEta <- 273.6 / (273.6+853.3) # 0.24
PTSD_parEta
mdd_parEta <- 573.4 / (573.4 + 853.3) # 0.4
mdd_parEta
```

- Test only depression to compare models

```{r}
model5_1 <- aov(EN ~ scid_mdd_pm, data=df)
summary(model5_1)
eta5_1 <- etasq(model5_1, partial = T)
anova(model5, model5_1)
BIC(model5, model5_1)
```
- Advantage to the model including both MDD and PTSD

```{r}
# test bdi on entire population
df_valid_bdi_all <- df[!is.na(df$`Scr BDI`), ] # 64 with valid BDI scores
df_valid_bdi_all$bdi <- as.numeric(df_valid_bdi_all$`Scr BDI`)
df_valid_bdi_all$caps_noEN <- df_valid_bdi_all$`CAPS Total` -  df_valid_bdi_all$EN
model6 <- lm(EN ~ bdi + caps_noEN, data=df_valid_bdi_all)
summary(model6)

eta6 <- etasq(model6) # 9%
```

- Compare to model with just BDI

```{r}
model6.2 <- lm(EN ~ bdi, data=df_valid_bdi_all)
summary(model6.2)
eta6.2 <- etasq(model6.2)
BIC(model6, model6.2)
anova(model6, model6.2) # sig. difference, by some points...
```

- Compare with model of just CAPS

```{r}
model6.3 <- lm(EN ~ caps_noEN, data=df_valid_bdi_all)
summary(model6.3)
eta6.3 <- etasq(model6.3)
BIC(model6, model6.2, model6.3)
anova(model6.2, model6.3)#, model6.3) # sig. difference, by some points...
```

## Test with RCF data

```{r}
dfRCF <- read.csv('/home/or/rewardPTSD/ReconsolidationOfFea_DATA_2021-03-02_1737.csv')

table(dfRCF$caps_ptsdc)
table(dfRCF$scid_mdd_pm)
table(dfRCF$caps_ptsdc, dfRCF$scid_mdd_pm)
```

```{r}
dfBDI <- dplyr::select(dfRCF, starts_with('bdi'), -bdiii_complete)
bdiTotal <- rowSums(dfBDI)
dfRCF$bdi_total <- bdiTotal

dfRCF_calc <- dplyr::select(dfRCF, subject_id,scid_mdd_pm, caps_ptsdc, caps_c4_im, caps_c5_im, caps_c6_im, caps_c4_fm,
  caps_c5_fm, caps_c6_fm, capstotalwithouten, calculatedtotal)
dfRCF_calc <- na.omit(dfRCF_calc)
dfRCF_calc$EN <- dfRCF_calc$calculatedtotal - dfRCF_calc$capstotalwithouten
dfRCF_calc$MDD <- case_when(dfRCF_calc$scid_mdd_pm == 3 ~ "YES", TRUE ~ "No")
dfRCF_calc$PTSD <- case_when(dfRCF_calc$caps_ptsdc == 1 ~ "YES", TRUE ~ "No")

# add bdi
bdiDF <- dplyr::select(dfRCF, subject_id, bdi_total)
bdiDF <- na.omit(bdiDF)
hist(bdiDF$bdi_total)
summary(bdiDF$bdi_total)
dfRCF_calc <- merge(dfRCF_calc, bdiDF, by = 'subject_id', all.x = T)
```

Do descriptive

```{r}
table(dfRCF$dem_gender)
table(dfRCF_calc$MDD)
table(dfRCF_calc$PTSD)
table(dfRCF_calc$MDD, dfRCF_calc$PTSD)
```

Run analysis

```{r}
model1_RCF <- aov(EN ~ MDD * PTSD, data=dfRCF_calc)
summary(model1_RCF)

etasq(model1_RCF) # here pcl explains almost 9%
pdf('FigureS1.pdf')
ggplot(dfRCF_calc, aes(x=MDD, y = EN,  color = PTSD)) + geom_boxplot() + geom_jitter() + ylab("Emotional Numbing Score") + theme_minimal()
dev.off()
```

Test model with just mdd

```{r}
model1.2RCF <- aov(EN ~ MDD, data=dfRCF_calc)
summary(model1.2RCF)
etasq(model1.2RCF)
BIC(model1_RCF, model1.2RCF)
anova(model1_RCF, model1.2RCF)
```

- Now do linear regression

```{r}
model2.RCF <- lm(EN ~ capstotalwithouten * bdi_total, data = dfRCF_calc)
summary(model2.RCF)
etasq(model2.RCF)
```

Model with just BDI

```{r}
model2.2RCF <- lm(EN ~ bdi_total, data = dfRCF_calc)
summary(model2.2RCF)
etasq(model2.2RCF)

BIC(model2.RCF, model2.2RCF)
anova(model2.RCF, model2.2RCF)
```

# Relative Importance

```{r}
capsLabels <- c("caps_b1_fm", "caps_b1_im","caps_b2_fm","caps_b2_im", "caps_b3_fm","caps_b3_im", "caps_b4_fm", "caps_b4_im","caps_b5_fm","caps_b5_im","caps_c1_fm", "caps_c1_im","caps_c2_fm","caps_c2_im","caps_c3_fm","caps_c3_im", "caps_c4_fm", "caps_c4_im", "caps_c5_fm","caps_c5_im","caps_c6_fm", "caps_c6_im", "caps_c7_fm", "caps_c7_im", "caps_d1_fm" ,"caps_d1_im", "caps_d2_fm", "caps_d2_im","caps_d3_fm","caps_d3_im", "caps_d4_fm", "caps_d4_im", "caps_d5_fm","caps_d5_im")
dfRCF_relaimp <- dplyr::select(dfRCF, subject_id, capsLabels, -bdiii_complete )
dfRCF_relaimp <- na.omit(dfRCF_relaimp)
dfRCF_relaimp <- merge(dfRCF_relaimp, bdiDF)
# rename columns (easier for plotting later)
# need to combine intensity and frequency of each symptom
data2 <- dplyr::select(dfRCF_relaimp, subject_id, bdi_total)
data2$B1 <- dfRCF_relaimp$caps_b1_fm + dfRCF_relaimp$caps_b1_im
data2$B2 <- dfRCF_relaimp$caps_b2_fm + dfRCF_relaimp$caps_b2_im
data2$B3 <- dfRCF_relaimp$caps_b3_fm + dfRCF_relaimp$caps_b3_im
data2$B4 <- dfRCF_relaimp$caps_b4_fm + dfRCF_relaimp$caps_b4_im
data2$B5 <- dfRCF_relaimp$caps_b5_fm + dfRCF_relaimp$caps_b5_im
data2$C1 <- dfRCF_relaimp$caps_c1_fm + dfRCF_relaimp$caps_c1_im
data2$C2 <- dfRCF_relaimp$caps_c2_fm + dfRCF_relaimp$caps_c2_im
data2$C3 <- dfRCF_relaimp$caps_c1_fm + dfRCF_relaimp$caps_c3_im
data2$C4 <- dfRCF_relaimp$caps_c4_fm + dfRCF_relaimp$caps_c4_im
data2$C5 <- dfRCF_relaimp$caps_c5_fm + dfRCF_relaimp$caps_c5_im
data2$C6 <- dfRCF_relaimp$caps_c6_fm + dfRCF_relaimp$caps_c6_im
data2$C7 <- dfRCF_relaimp$caps_c7_fm + dfRCF_relaimp$caps_c7_im
data2$D1 <- dfRCF_relaimp$caps_d1_fm + dfRCF_relaimp$caps_d1_im
data2$D2 <- dfRCF_relaimp$caps_d2_fm + dfRCF_relaimp$caps_d2_im
data2$D3 <- dfRCF_relaimp$caps_d3_fm + dfRCF_relaimp$caps_d3_im
data2$D4 <- dfRCF_relaimp$caps_d4_fm + dfRCF_relaimp$caps_d4_im
data2$D5 <- dfRCF_relaimp$caps_d5_fm + dfRCF_relaimp$caps_d5_im
data2$EN <- data2$C4 + data2$C5 + data2$C6
```

Model
```{r}
model_all_data2 <- lm(EN ~ B1 + B2 + B3 + B4 + B5 + C1 + C2 + C3 + 
                  C7 + D1 + D2 + D3 + D4 + D5 + bdi_total, data = data2)
summary(model_all_data2)
calc.relimp(model_all_data2, type="lmg")
```


```{r}
boot2 <- boot.relimp(model_all_data2, type = 'lmg', b = 1000)

booteval2 <- booteval.relimp(boot2)
plot(booteval2)
```
## generate nicer graph

```{r}
label = c('B1','B2','B3','B4','B5','C1','C2', 'C3', 'C7', 'D1','D2','D3','D4','D5', 'BDI')
lgmValues <- booteval2@lmg # mean of each variable
lgmLow <- booteval2@lmg.lower[1,] # create as vector instead of 1D matrix
lgmHigh <- booteval2@lmg.upper[1,]

lmgData2 <- tibble(lgmValues, lgmLow, lgmHigh, label)
# sort data by value
lmgOrdered2 <- lmgData2[order(lgmValues, decreasing = F),]
# set label as factor with ordered levels
lmgOrdered2$label <- factor(lmgOrdered2$label, levels = lmgOrdered2$label)
#plot
pdf('Figure_S2.pdf')
ggplot(lmgOrdered2, aes(x=label, y=(lgmValues*100))) +  geom_bar(stat="identity", fill='lightblue',
           position=position_dodge()) +  geom_errorbar(aes(ymin=(lgmLow*100), ymax=(lgmHigh*100)), width=.2,
                 position=position_dodge(.9)) + coord_flip() + ylab("Values [LMG in %]") + xlab('PTSD Symptom')  +theme_minimal()
dev.off()
```

## Difference graph

```{r}

diffLabels <- c("B1_B2", "B1_B3","B1_B4", "B1_B5", "B1_C1", "B1_C2", "B1_C3","B1_C7", "B1_D1", "B1_D2", "B1_D3",
              "B1_D4", "B1_D5", "B1_BDI", "B2_B3","B2_B4","B2_B5", "B2_C1","B2_C2","B2_C3","B2_C7", "B2_D1", "B2_D2", "B2_D3","B2_D4", "B2_D5", "B2_BDI", "B3_B4","B3_B5", "B3_C1", "B3_C2", "B3_C3","B3_C7", "B3_D1", "B3_D2", "B3_D3", "B3_D4", "B3_D5", "B3_BDI", "B4_B5", "B4_C1", "B4_C2", "B4_C3","B4_C7", "B4_D1", "B4_D2", "B4_D3", "B4_D4", "B4_D5", "B4_BDI", "B5_C1", "B5_C2", "B5_C3", "B5_C7", "B5_D1","B5_D2", "B5_D3", "B5_D4", "B5_D5", "B5_BDI", "C1_C2","C1_C3","C1_C7","C1_D1", "C1_D2", "C1_D3", "C1_D4", "C1_D5","C1_BDI", "C2_C3","C2_C7", "C2_D1","C2_D2","C2_D3", "C2_D4", "C2_D5","C2_BDI", "C3_C7", "C3_D1", "C3_D2","C3_D3","C3_D4","C3_D5","C3_BDI", "C7_D1","C7_D2","C7_D3","C7_D4","C7_D5","C7_BDI", "D1_D2","D1_D3", "D1_D4", "D1_D5", "D1_BDI", "D2_D3", "D2_D4", "D2_D5", "D2_BDI", "D3_D4", "D3_D5", "D3_BDI", "D4_D5", "D4_BDI", "D5_BDI")
dfDiff2 <- booteval2@lmg.diff
dfDiff_low2 <- booteval2@lmg.diff.lower[1,]
dfDiff_high2 <- booteval2@lmg.diff.upper[1,]
dfDiff2 <- tibble(dfDiff2, dfDiff_high2, dfDiff_low2, diffLabels)

#dfDiff <- booteval@markdiff
pdf("diffGraph_data2.pdf", width = 10, height = 20)
#png("diffGraph.png", width = 800, height = 1024)
ggplot(dfDiff2, aes(x=diffLabels, y=dfDiff2)) +  geom_bar(stat="identity",
           position=position_dodge()) +  geom_errorbar(aes(ymin=dfDiff_low2, ymax=dfDiff_high2), width=.2,
                 position=position_dodge(.9)) + coord_flip() + ylab("Difference in Values [LMG]") + xlab('PTSD Symptom pair')  +theme_minimal()
dev.off()
```

```{r}
dfRCF_age <- read.csv('/home/or/rewardPTSD/ReconsolidationOfFea_DATA_2021-03-03_0913.csv')
na.omit(dfRCF_age)
summary(dfRCF_age$dem_age)
sd(dfRCF_age$dem_age)
```
















