---
title: "Anhedonia in PTSD and MDD"
author: "Or Duek"
output:
  html_document:
    df_print: paged
---

## Assessing anhedonia in PTSD

```{r, echo=FALSE, results='hide', comment=FALSE, warning=FALSE, message=FALSE}
require(tidyverse)
require(readxl)
require(ggplot2)
require(lme4)
require(lmerTest)
require(ppcor)
require(lsr)
require(heplots)
```

```{r, echo=FALSE}
# load data
df <- read_excel('/home/or/PTSD_reward_data/Cleaned_Official_Data_Set_11112020.xlsx')
```


```{r, echo=FALSE}
ggplot(df, aes(y=`Crit D Total`, x= scid_mdd_pm, color=`CAPS PTSD`)) + geom_boxplot() + 
  theme_minimal()
```

* This shows nicely that PTSD patients score higher in the anhedonia scores, even when not depressed
* Lets tests it using anova

```{r}
model1 <- aov(`Crit D Total` ~ scid_mdd_pm*`CAPS PTSD`, data=df)
summary(model1)

#eta1 <- etasq(model1, partial = FALSE) # don't trust this function
# calculate manually
#model1

PTSD_parEta <- 1581.2754 / (1581.2754+2367.4913) # 0.4
PTSD_parEta
mdd_parEta <- 2044.7645 / (2044.7645 + 2367.4913) # 0.46
mdd_parEta
```

- As expected, we see a significant effect for both MDD and PTSD.
- 40% partial eta square for the PTSD diagnosis, while only 31% for depression. (in different calculation we have 46% explained by depression). 
- Lets use PCL (which is continuous and we don't use the exact same measure)


```{r}
# using the total CAPS score
model2 <- aov(`Crit D Total` ~ scid_mdd_pm +`Scr PCL`, data=df)
summary(model2)


```


- PCL scores are not significant, which might be problematic (as we can claim the main reason for D expaind by PTSD is that they use same measure)
- We lack enough BDI scores to use a covariate (which I think will be the best option)







